---
title: "HDChangePoint"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set( collapse = TRUE,
  comment = "#>",
  fig.path = "README-")
```

## Overview
The R package `HDChangePoint` implements two estimation methods using the parametric nonlinear mixed effects model (NLME) and the multi-stage nonparametric approaches, which are described in the following manuscript: 
  
  **U.Lee, R.J.Carroll, K.Marder, Y.wang and T.P.Garcia. (2019+) "Estimating Disease Onset from Change Points of Markers Measured with Error"**
  
  Those two methods estimate individual longitudinal trajectories in an "S"-shape and their respective inflection points.


## Installation 

To install `HDChangePoint` from GitHub,
```{r, eval=FALSE}
devtools::install_github("unkyunglee/HDChangePoint")
```

## Usage
```{r usage, eval=F, message=FALSE}

library(HDChangePoint)

# Specify parameters for data generation
n=80;
model="logist";
p=2;
bb0=0.5;
bb=0.1;
x.sd=0.3;
v1=5;
v2=7;
dist="normal";
eps.sd=0.05;
u.sd=0.05;

set.seed(22)
# Generate a dataset under the logistic model
outdat<-mydata(n=n, model=model, p=p, bb0=bb0, bb=bb, x.sd=x.sd, 
               v1=v1, v2=v2, dist=dist, eps.sd=eps.sd, u.sd=u.sd)


# Specify other parameters for the multi-stage nonparametric procedure
num.interp=45;
newl=45;
k1=20;
k2=20;
time.length=20;
mean.diff=1;
tolerance=0.009
iter=0;

# Multi-stage nonparametric estimation
results<-sim.nonpara(n=n, model="logist", dist="normal", k1=k1, k2=k2, num.interp=num.interp, 
                     newl=newl, eps.sd=eps.sd, mean.diff=1, tolerance=tolerance,
                     itermax=50, iter=iter, time.length=time.length, dat=outdat)

```

## Example 1.

We give an example to reproduce the part of the simulation results in Table 1 in our manuscript. We consider the `nonpara_logist_visit1` data available from R package `HDChangePoint`. It contains simluation results from the multi-stage nonparametric estimation when the true data are generated from the logistic model with the total number of visit numbers $m=5,6,$ or $7$. 

```{r ex1, message=FALSE}
library(HDChangePoint)
data(nonpara_logist_visit1_nsim500)
```

We specify parameters and obtain the results as the part of the Table 1 and plots (c) and (d) in Figure 2. Figures will be automatrically saved as .eps files in your working directory. 

```{r simuout, cache=TRUE}
# Specify parameters in the summary function
nsim=500;
n=80;
num.interp=45;
newl=45;
time.length=20;
bb.true=0.1;
b0.true=0.5;
true.sigma.u=0.05;
true.sigma.eps=0.05;

# Produce the results 
simus.out1<-simus.results(nsim=nsim, n=n, res=nonpara_logist_visit1_nsim500,
                          model="logist", num.interp=num.interp, newl=newl,
                          time.length=time.length,
                          bb.true=bb.true, b0.true=b0.true, 
                          true.sigma.u=true.sigma.u, true.sigma.eps=true.sigma.eps,
                          file="logist_visit1")




simus.out1$res.table
simus.out1$relative.avr.est.Z
```

## Example 2. 

We give an example to show reproducibility for data analysis study in our manuscript. We consider the `PSEUDO_PREDICT_HD` data available from R package `HDChangePoint`. This data is a simulated dataset using the data generation function `pseudo.HD()`, see `data-raw\PSEUDO_PREDICT_HD.R`. 

```{r pseudo_data, results='hide'}
library(HDChangePoint)
data(PSEUDO_PREDICT_HD)
```



The data consist of 80 subjects' information with 8 variables. Each subject made different number of visits $m=5,6$ or $7$ at clinics. In the `PSEUDO_PREDICT_HD` data, we consider two subject specific covariates: CAG repeats and gender, which may be associated with inflection points.

```{r data_info, message=FALSE}
head(PSEUDO_PREDICT_HD)
?PSEUDO_PREDICT_HD # this gives you more information on the dataset
```

We fit our methods to the `PSEUDO_PREDICT_HD` data. First, we specify the parameters and run the function `hd.study()`.
  
```{r pseudo_analysis, message=FALSE, results='hide'}

# Specify the parameters
simu.dat<-PSEUDO_PREDICT_HD
n=80;
m=42;
num.interp=42;
newl=42;
mean.diff=1;
tolerance=0.01;
itermax=20;
iter=0;

# produce two tables to reproduce similar results as in Table 4 and Figure 1 of our manuscript
simu.analysis.results<-hd.study(simu.data=simu.data, m=m, num.interp=num.interp, 
                                n=n, newl=newl, mean.diff=mean.diff, 
                                tolerance=tolerance, itermax=itermax, iter=iter,     
                                boot.ci=TRUE)

```

We obtain the following results for the multi-stage nonparametric estimates and the parametric NLME estimates, similar to Table 4 in our manuscript. Figure will be automatically saved in your working directory, which is similar to Figure 1 in our manuscript. You can also find the result figure in the following folder: man \\ examples\_figures
```{r analysis_results}
# produce results
simu.analysis.results$nonpara_summary_table4 # multi-stage nonparametric estimates
simu.analysis.results$para_summary_table4    # parametric NLME estimates
```


